Non-word repetition in Yélî Dnye

### 2021-01-05 manuscript.Rmd
stopped midway the results

We then analyzed variation in whole-item NWR scores as a function of the average frequency with which sounds composing individual target words are found in languages over the world. To look at this, we fit a mixed logistic regression in which the outcome variable was whether the non-word was correctly repeated or not. The fixed effect of interest was the average cross-linguistic phone frequency; we also included child age as a control fixed effect, and child ID & target ID as random variables.

@@

We could include `r summary(xl_fr)$devcomp$dims["N"]` observations, from `r summary(xl_fr)$ngrps["file"]` children producing in any given trial one of `r summary(xl_fr)$ngrps["target"]` potential target words. The analysis revealed a main effect of age (ß = `r mycoeff["age.rounded_s","Estimate"]`, SE ß = `r mycoeff["age.rounded_s","Std. Error"]`, p `r pval.display(mycoeff["age.rounded_s","Pr(>|z|)"])`); and a significant estimate for the scaled average cross-linguistic frequency of phones in the target words (ß = `r mycoeff["avg_fr_s","Estimate"]`, SE ß = `r mycoeff["avg_fr_s","Std. Error"]`, p `r pval.display(mycoeff["avg_fr_s","Pr(>|z|)"])`): Target words with phones found more frequently across languages had a higher proportion of words that were correctly repeated, as shown in Figure \@ref(fig:Fig2-xling-freq). Averaging across items and participants, the Pearson correlation between scaled average cross-linguistic phone frequency and whole-item NWR scores was r(`r xlfr_r$parameter`) = `r round(xlfr_r$estimate,3)`.

@@

We next checked whether the association between whole-item NWR scores and cross-linguistic phone frequence could actually be due to frequency of the sounds within the language: One can suppose that sounds that occur more frequently across languages are also more frequent within a language, and therefore may be easier for children to represent and repeat. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings [@casillas2020early] by counting the number of word types in which they occurred, and applied the natural logarithm.[^2] Here, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Phone frequencies estimated from this corpus were correlated with cross-linguistic phone frequencies (r(`r freq_cor$parameter`)=`r round(freq_cor$estimate,3)`, p `r pval.display(round(freq_cor$p.value,3))`). Moreover, averaging across items and participants, the Pearson correlation between scaled average corpus phone frequency and whole-item NWR scores was r = `r round(cor.test(cor_fr_tab$corf,cor_fr_tab$Word)$estimate,3)`. Therefore, we fit another mixed logistic regression, this time declaring as fixed effects both scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic frequency (ß = `r both_coeff["avg_fr_s","Estimate"]`, SE ß = `r both_coeff["avg_fr_s","Std. Error"]`, p `r pval.display(both_coeff["avg_fr_s","Pr(>|z|)"])`) and age (ß = `r both_coeff["age.rounded_s","Estimate"]`, SE ß = `r both_coeff["age.rounded_s","Std. Error"]`, p `r pval.display(both_coeff["age.rounded_s","Pr(>|z|)"])`) were significant predictors of whole-item  NWR scores, but corpus frequency (ß = `r both_coeff["avg_fr_cor_ty_s","Estimate"]`, SE ß = `r both_coeff["avg_fr_cor_ty_s","Std. Error"]`, p `r pval.display(both_coeff["avg_fr_cor_ty_s","Pr(>|z|)"])`) was not.

@@

sub_tab=table(sub_bank)[allVC,allVC]

@@

Next, we investigated patterns of deletion and substitution (as insertion was not attested), this time looking at all attempts so as to base our generalizations on more data. Deletions were relatively rare: there were only `r sum(del_tab[c(oralVs,nasalVs)])` instances of deleted vowels (~`r mean(del_pc[c(oralVs,nasalVs)],na.rm=T)`% of all vowel targets), and `r sum(del_tab[c(simplexCs,complexCs)])` instances of deleted consonants (~`r mean(del_pc[c(simplexCs,complexCs)],na.rm=T)`% of all consonant targets). Substitutions were somewhat more frequent, but still relatively rare. Nasal vowels were produced as oral vowels and vice versa (`r sum(sub_tab[oralVs,nasalVs])` oral targets as nasal (~`r mean(sub_pc[oralVs,nasalVs],na.rm=T)`% of targets) and `r sum(sub_tab[nasalVs,oralVs])` nasal targets produced as oral (~`r mean(sub_pc[nasalVs,oralVs],na.rm=T)`% of targets)). Both oral and nasal vowels were also occasionally subsituted by vowels of a different quality that preserved the target oral/nature airflow (`r sum(sub_tab[oralVs,oralVs])` oral vowels (~`r mean(sub_pc[oralVs,oralVs],na.rm=T)`% of targets and `r sum(sub_tab[nasalVs,nasalVs])` nasal vowels (~`r mean(sub_pc[nasalVs,nasalVs],na.rm=T)`% of targets). Among consonants, complex consonants were occasionally substituted by simpler ones (specifically `r print.phone.list(complexCs.ipa)` as `r print.phone.list(simplexCs.ipa)`), `r sum(sub_tab[complexCs,simplexCs])` times, in approximately `r mean(sub_pc[complexCs,simplexCs],na.rm=T)`% of all complex consonant targets. Interestingly, simple consonants were sometimes also substituted by complex ones; this happened `r sum(sub_tab[simplexCs,complexCs])` times, in approximately `r mean(sub_pc[complexCs,simplexCs],na.rm=T)`% of all simple consonant targets. Finally, simple consonants were mispronounced as other simple consonants (`r sum(sub_tab[simplexCs,simplexCs])` instances, ~`r mean(sub_pc[simplexCs,simplexCs],na.rm=T)`% of all simple consonant targets) and complex consonants as other complex consonants (`r sum(sub_tab[complexCs,complexCs])` instances, ~`r mean(sub_pc[complexCs,complexCs],na.rm=T)`% of all complex consonant targets). Note that the substitutions included phones that are not native to Yélî Dnye but do occur in English (e.g., [tʃ]); as these data come from careful transcriptions by a native Yélî Dnye speaker who is very fluent in English, we take these segmental substitutions faithfully as an indication that several of our participants have mastered production of some English phones, possibly produced within whole English word forms.

Finally, we looked into the frequency with which mispronunciations resulted in real words. In fact, two thirds of incorrect repetitions were recognizable as real words or phrases in Yélî Dnye or English: `r round(sum(!is.na(trnsc.inc$meaning) & trnsc.inc$meaning != "0")/sum(!is.na(trnsc.inc$mispronunciation))*100)`%. This type of analysis is seldom reported---we could only find one comparison point: @castro1998illiterate found that illiterate European Portuguese adults' NWR mispronunciations resulted in real words in `r round(53/475*100,2)`% of cases, whereas literate participants did so in only `r round(2/117*100,2)`% of cases. The percentage we observe here is much higher than reported in Castro and colleagues' study, but we do not know whether age, language, or even test structure explains this difference.

@@

Next, we inspected whether NWR scores varied as a function of word length (Table \@ref(tab:tablength)). Participants scored much lower on monosyllables than on non-words of other lengths. This is likely due to the fact that the majority of monosyllables were designed to include sounds that are rare in the world's languages, which may indicate that they are hard to produce or perceive. Setting monosyllables aside, we observe the typical pattern of lower scores for longer items, although this is particularly salient for the whole-item scoring. While whole-item scoring is the most commonly reported NWR outcome, it is also the least forgiving. The pattern is less marked when other two scores are used, which are less sensitive to errors. Averaging across participants and items, the Pearson correlation between length (2-4 syllables) and whole-item NWR scores was r(`r length_r$parameter`) = `r round(length_r$estimate,3)`. In a generalized binomial mixed model, we included `r summary(len_lm)$devcomp$dims["N"]` observations, from `r summary(len_lm)$ngrps["file"]` children producing, in any given trial, one of `r summary(len_lm)$ngrps["target"]` potential target words. The analysis revealed a positive effect of age (ß = `r mycoeff["age.rounded_s","Estimate"]`, SE ß = `r mycoeff["age.rounded_s","Std. Error"]`, p `r pval.display(mycoeff["age.rounded_s","Pr(>|z|)"])`) and a negative but non-significant estimate for target length in number of syllables (ß = `r mycoeff["type","Estimate"]`, SE ß = `r mycoeff["type","Std. Error"]`, p `r pval.display(mycoeff["type","Pr(>|z|)"])`).

```{r tablength, results = "asis"}
length.tab=NULL
for(myvar in c("Word","Phoneme","NLD")){
  length.tab=cbind(length.tab,
              paste0(round(aggregate(by_length[,myvar],by=list(by_length$Length),mean)$x),
              " (",round(aggregate(by_length[,myvar],by=list(by_length$Length),sd)$x),")"))
}
rownames(length.tab)<-levels(by_length$Length)
colnames(length.tab)<-c("Word","Phoneme","NLD")
apa_table(length.tab,caption="NWR measured in whole-word scores, phoneme-based scores, and normalized Levenshtein Distance, separately for the four stimuli lengths.")
```

>>>>

We then analyzed variation in whole-item NWR scores as a function of the average frequency with which sounds composing individual target words are found in languages over the world. To look at this, we fit a mixed logistic regression in which the outcome variable was whether the non-word was correctly repeated or not. The fixed effect of interest was the average cross-linguistic phone frequency; we also included child age as a control fixed effect, and allowed slopes to vary over the random effects child ID and target ID.

@@

We could include `r summary(xl_fr)$devcomp$dims["N"]` observations, from `r summary(xl_fr)$ngrps["file"]` children producing in any given trial one of `r summary(xl_fr)$ngrps["target"]` potential target words. The analysis revealed a main effect of age (ß = `r mycoeff["age.rounded_s","Estimate"]`, SE ß = `r mycoeff["age.rounded_s","Std. Error"]`, p `r pval.display(mycoeff["age.rounded_s","Pr(>|z|)"])`); and a significant estimate for the scaled average cross-linguistic frequency of phones in the target words (ß = `r mycoeff["avg_fr_s","Estimate"]`, SE ß = `r mycoeff["avg_fr_s","Std. Error"]`, p `r pval.display(mycoeff["avg_fr_s","Pr(>|z|)"])`): Target words with phones found more frequently across languages had higher correct repetition scores, as shown in Figure \@ref(fig:Fig2-xling-freq). Averaging across items and participants, the Pearson correlation between scaled average cross-linguistic phone frequency and whole-item NWR scores was r(`r xlfr_r$parameter`) = `r round(xlfr_r$estimate,3)`.

@@

We next checked whether the association between whole-item NWR scores and cross-linguistic phone frequence could actually be due to frequency of the sounds within the language: One can suppose that sounds that occur more frequently across languages are also more frequent within a language, and therefore may be easier for children to represent and repeat. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings [@casillas2020early] by counting the number of word types in which they occurred, and applied the natural logarithm.[^2] Here, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Item-level average corpus-based phone frequencies were correlated with the item-level average cross-linguistic phone frequencies (r(`r freq_cor$parameter`)=`r round(freq_cor$estimate,3)`, p `r pval.display(round(freq_cor$p.value,3))`). Moreover, averaging across items and participants, the Pearson correlation between scaled average corpus phone frequency and whole-item NWR scores was r = `r round(cor.test(cor_fr_tab$corf,cor_fr_tab$Word)$estimate,3)`. Therefore, we fit another mixed logistic regression, this time declaring as fixed effects both scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic phone frequency (ß = `r both_coeff["avg_fr_s","Estimate"]`, SE ß = `r both_coeff["avg_fr_s","Std. Error"]`, p `r pval.display(both_coeff["avg_fr_s","Pr(>|z|)"])`) and age (ß = `r both_coeff["age.rounded_s","Estimate"]`, SE ß = `r both_coeff["age.rounded_s","Std. Error"]`, p `r pval.display(both_coeff["age.rounded_s","Pr(>|z|)"])`) were significant predictors of whole-item  NWR scores, but corpus phone frequency (ß = `r both_coeff["avg_fr_cor_ty_s","Estimate"]`, SE ß = `r both_coeff["avg_fr_cor_ty_s","Std. Error"]`, p `r pval.display(both_coeff["avg_fr_cor_ty_s","Pr(>|z|)"])`) was not.

@@

sub_tab=table(sub_bank)[allVC,allVC]
#this table has target as rows, realization as columns

@@

Next, we investigated patterns of deletion and substitution (as insertion was not attested), this time looking at all attempts so as to base our generalizations on more data. Deletions were very rare: there were only `r sum(del_tab[c(oralVs,nasalVs)])` instances of deleted vowels (~`r mean(del_pc[c(oralVs,nasalVs)],na.rm=T)`% of all vowel targets), and `r sum(del_tab[c(simplexCs,complexCs)])` instances of deleted consonants (~`r mean(del_pc[c(simplexCs,complexCs)],na.rm=T)`% of all consonant targets). We therefore focus our qualitative description here on substitutions: There were `r dim(sub_bank)[1]` cases of substitutions, ~`r round(dim(sub_bank)[1]/(dim(del_bank)[1] + dim(sub_bank)[1] + dim(match_bank)[1])*100,2)` of the `r dim(del_bank)[1] + dim(sub_bank)[1] + dim(match_bank)[1]` phones found collapsing across all children and target words, so that substitutions constituted the frank majority of incorrect phones (~`r round(dim(sub_bank)[1]/(dim(del_bank)[1] + dim(sub_bank)[1] )*100,2)` of unmatched phones).

<!-- reminder: this table has target as rows, realization as columns  -->
Is it the case that cross-linguistically less common phones are more frequently substituted by more common ones than vice versa? We found that errors involving nasal vowel targets tended to result in a change in nasality (`r sum(sub_tab[nasalVs,oralVs])` nasal targets produced as oral, ~`r round(sum(sub_tab[nasalVs,oralVs])/sum(sub_tab[nasalVs,])*100)`% of substitutions for nasal vowel target phones) more frequently than in a change in quality (`r sum(sub_tab[nasalVs,nasalVs])` nasal targets produced as a different nasal vowel, ~`r round(sum(sub_tab[nasalVs,nasalVs])/sum(sub_tab[nasalVs,])*100)`% of substitutions for nasal targets). The opposite was true for oral vowels: there were `r sum(sub_tab[oralVs,nasalVs])` oral targets produced as nasal, constituting ~`r round(sum(sub_tab[oralVs,nasalVs])/sum(sub_tab[oralVs,])*100)`% of substitutions for oral target phones; compared to `r sum(sub_tab[oralVs,oralVs])` oral targets produced as a different oral vowel, ~`r round(sum(sub_tab[oralVs,oralVs])/sum(sub_tab[oralVs,])*100)`% of substitutions for oral targets).

**STOPPED HERE**

Among consonants, complex consonants were occasionally substituted by simpler ones (specifically `r print.phone.list(complexCs.ipa)` as `r print.phone.list(simplexCs.ipa)`), `r sum(sub_tab[complexCs,simplexCs])` times, in approximately `r mean(sub_pc[complexCs,simplexCs],na.rm=T)`% of all complex consonant targets. Interestingly, simple consonants were sometimes also substituted by complex ones; this happened `r sum(sub_tab[simplexCs,complexCs])` times, in approximately `r mean(sub_pc[complexCs,simplexCs],na.rm=T)`% of all simple consonant targets. Finally, simple consonants were mispronounced as other simple consonants (`r sum(sub_tab[simplexCs,simplexCs])` instances, ~`r mean(sub_pc[simplexCs,simplexCs],na.rm=T)`% of all simple consonant targets) and complex consonants as other complex consonants (`r sum(sub_tab[complexCs,complexCs])` instances, ~`r mean(sub_pc[complexCs,complexCs],na.rm=T)`% of all complex consonant targets). Note that the substitutions included phones that are not native to Yélî Dnye but do occur in English (e.g., [tʃ]); as these data come from careful transcriptions by a native Yélî Dnye speaker who is very fluent in English, we take these segmental substitutions faithfully as an indication that several of our participants have mastered production of some English phones, possibly produced within whole English word forms.

Finally, we looked into the frequency with which mispronunciations resulted in real words. In fact, two thirds of incorrect repetitions were recognizable as real words or phrases in Yélî Dnye or English: `r round(sum(!is.na(trnsc.inc$meaning) & trnsc.inc$meaning != "0")/sum(!is.na(trnsc.inc$mispronunciation))*100)`%. This type of analysis is seldom reported. We could only find one comparison point: @castro1998illiterate found that illiterate European Portuguese adults' NWR mispronunciations resulted in real words in `r round(53/475*100,2)`% of cases, whereas literate participants did so in only `r round(2/117*100,2)`% of cases. The percentage we observe here is much higher than reported in Castro and colleagues' study, but we do not know whether age, language, or even test structure explains this difference.

@@

Next, we inspected whether NWR scores varied as a function of word length (Table \@ref(tab:tablength)). Participants scored much lower on monosyllables than on non-words of other lengths. This is likely due to the fact that the majority of monosyllables were designed to include sounds that are rare in the world's languages, which may indicate that they are hard to produce or perceive. Setting monosyllables aside, we observe the typical pattern of lower scores for longer items, although this is particularly salient for the whole-item scoring. While whole-item scoring is the most commonly reported NWR outcome, it is also the least forgiving. Averaging across participants and items, the Pearson correlation between length (2-4 syllables) and whole-item NWR scores was r(`r length_r$parameter`) = `r round(length_r$estimate,3)`. In a generalized binomial mixed model, we included `r summary(len_lm)$devcomp$dims["N"]` observations, from `r summary(len_lm)$ngrps["file"]` children producing, in any given trial, one of `r summary(len_lm)$ngrps["target"]` potential target words. The analysis revealed a positive effect of age (ß = `r mycoeff["age.rounded_s","Estimate"]`, SE ß = `r mycoeff["age.rounded_s","Std. Error"]`, p `r pval.display(mycoeff["age.rounded_s","Pr(>|z|)"])`) and a negative but non-significant estimate for target length in number of syllables (ß = `r mycoeff["type","Estimate"]`, SE ß = `r mycoeff["type","Std. Error"]`, p `r pval.display(mycoeff["type","Pr(>|z|)"])`).

```{r tablength, results = "asis"}
length.tab=NULL
for(myvar in c("Word","NLD")){
  length.tab=cbind(length.tab,
              paste0(round(aggregate(by_length[,myvar],by=list(by_length$Length),mean)$x),
              " (",round(aggregate(by_length[,myvar],by=list(by_length$Length),sd)$x),")"))
}
rownames(length.tab)<-levels(by_length$Length)
colnames(length.tab)<-c("Word","NLD")
apa_table(length.tab,caption="NWR measured in whole-word scores, and normalized Levenshtein Distance, separately for the four stimuli lengths.")
```

### 2021-01-05 manuscript.tex
stopped midway the results

We then analyzed variation in whole-item NWR scores as a function of the average frequency with which sounds composing individual target words are found in languages over the world. To look at this, we fit a mixed logistic regression in which the outcome variable was whether the non-word was correctly repeated or not. The fixed effect of interest was the average cross-linguistic phone frequency; we also included child age as a control fixed effect, and child ID \& target ID as random variables.

We could include 850 observations, from 40 children producing in any given trial one of 41 potential target words. The analysis revealed a main effect of age (ß = 0.33, SE ß = 0.13, p = 0.01); and a significant estimate for the scaled average cross-linguistic frequency of phones in the target words (ß = 0.82, SE ß = 0.18, p \textless{} 0.001): Target words with phones found more frequently across languages had a higher proportion of words that were correctly repeated, as shown in Figure \ref{fig:Fig2-xling-freq}. Averaging across items and participants, the Pearson correlation between scaled average cross-linguistic phone frequency and whole-item NWR scores was r(32) = 0.61.

@@

We next checked whether the association between whole-item NWR scores and cross-linguistic phone frequence could actually be due to frequency of the sounds within the language: One can suppose that sounds that occur more frequently across languages are also more frequent within a language, and therefore may be easier for children to represent and repeat. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings (Casillas et al., 2020) by counting the number of word types in which they occurred, and applied the natural logarithm.\footnote{We also carried out this analysis using token phone frequency, but this measure was not correlated with whole-item NWR scores, and therefore the fact that it did not explain away the predictive value of cross-linguistic phone frequency was less informative than the relationship discussed in the main text, with type frequencies.} Here, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Phone frequencies estimated from this corpus were correlated with cross-linguistic phone frequencies (r(27)=0.50, p \textless{} 0.01). Moreover, averaging across items and participants, the Pearson correlation between scaled average corpus phone frequency and whole-item NWR scores was r = 0.46. Therefore, we fit another mixed logistic regression, this time declaring as fixed effects both scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic frequency (ß = 0.82, SE ß = 0.27, p \textless{} 0.01) and age (ß = 0.33, SE ß = 0.13, p = 0.01) were significant predictors of whole-item NWR scores, but corpus frequency (ß = 0.01, SE ß = 0.25, p = 0.98) was not.

@@

Next, we investigated patterns of deletion and substitution (as insertion was not attested), this time looking at all attempts so as to base our generalizations on more data. Deletions were relatively rare: there were only 12 instances of deleted vowels (\textasciitilde{}0.28\% of all vowel targets), and 6 instances of deleted consonants (\textasciitilde{}0.19\% of all consonant targets). Substitutions were somewhat more frequent, but still relatively rare. Nasal vowels were produced as oral vowels and vice versa (52 oral targets as nasal (\textasciitilde{}1.10\% of targets) and 58 nasal targets produced as oral (\textasciitilde{}2.53\% of targets)). Both oral and nasal vowels were also occasionally subsituted by vowels of a different quality that preserved the target oral/nature airflow (205 oral vowels (\textasciitilde{}2.16\% of targets and 23 nasal vowels (\textasciitilde{}2.36\% of targets). Among consonants, complex consonants were occasionally substituted by simpler ones (specifically {[}ṭp{]}, {[}tp{]}, {[}kp{]}, {[}km{]}, {[}kṇ{]}, {[}mp{]}, {[}ɣ{]}, and {[}lβʲ{]} as {[}m{]}, {[}n{]}, {[}l{]}, {[}w{]}, {[}j{]}, {[}w{]}, {[}ṭ{]}, {[}g{]}, {[}p{]}, {[}t{]}, {[}k{]}, {[}f{]}, {[}h{]}, and {[}tʃ{]}), 250 times, in approximately 2.43\% of all complex consonant targets. Interestingly, simple consonants were sometimes also substituted by complex ones; this happened 2 times, in approximately 2.43\% of all simple consonant targets. Finally, simple consonants were mispronounced as other simple consonants (120 instances, \textasciitilde{}0.52\% of all simple consonant targets) and complex consonants as other complex consonants (55 instances, \textasciitilde{}0.98\% of all complex consonant targets). Note that the substitutions included phones that are not native to Yélî Dnye but do occur in English (e.g., {[}tʃ{]}); as these data come from careful transcriptions by a native Yélî Dnye speaker who is very fluent in English, we take these segmental substitutions faithfully as an indication that several of our participants have mastered production of some English phones, possibly produced within whole English word forms.

Finally, we looked into the frequency with which mispronunciations resulted in real words. In fact, two thirds of incorrect repetitions were recognizable as real words or phrases in Yélî Dnye or English: 63\%. This type of analysis is seldom reported---we could only find one comparison point: Castro-Caldas, Petersson, Reis, Stone-Elander, and Ingvar (1998) found that illiterate European Portuguese adults' NWR mispronunciations resulted in real words in 11.16\% of cases, whereas literate participants did so in only 1.71\% of cases. The percentage we observe here is much higher than reported in Castro and colleagues' study, but we do not know whether age, language, or even test structure explains this difference.

@@

Next, we inspected whether NWR scores varied as a function of word length (Table \ref{tab:tablength}). Participants scored much lower on monosyllables than on non-words of other lengths. This is likely due to the fact that the majority of monosyllables were designed to include sounds that are rare in the world's languages, which may indicate that they are hard to produce or perceive. Setting monosyllables aside, we observe the typical pattern of lower scores for longer items, although this is particularly salient for the whole-item scoring. While whole-item scoring is the most commonly reported NWR outcome, it is also the least forgiving. The pattern is less marked when other two scores are used, which are less sensitive to errors. Averaging across participants and items, the Pearson correlation between length (2-4 syllables) and whole-item NWR scores was r(1) = -0.91. In a generalized binomial mixed model, we included 479 observations, from 40 children producing, in any given trial, one of 24 potential target words. The analysis revealed a positive effect of age (ß = 0.56, SE ß = 0.14, p \textless{} 0.001) and a negative but non-significant estimate for target length in number of syllables (ß = -0.15, SE ß = 0.33, p = 0.65).

@@

\caption{\label{tab:tablength}NWR measured in whole-word scores, phoneme-based scores, and normalized Levenshtein Distance, separately for the four stimuli lengths.}

\begin{tabular}{lll}
\toprule
Word & \multicolumn{1}{c}{Phoneme} & \multicolumn{1}{c}{NLD}\\
\midrule
47 (22) & 59 (17) & 41 (17)\\
79 (22) & 92 (9) & 8 (9)\\
78 (19) & 93 (7) & 7 (7)\\
74 (32) & 91 (12) & 9 (12)\\
\bottomrule
\end{tabular}

>>>>

We then analyzed variation in whole-item NWR scores as a function of the average frequency with which sounds composing individual target words are found in languages over the world. To look at this, we fit a mixed logistic regression in which the outcome variable was whether the non-word was correctly repeated or not. The fixed effect of interest was the average cross-linguistic phone frequency; we also included child age as a control fixed effect, and allowed slopes to vary over the random effects child ID and target ID.

We could include 850 observations, from 40 children producing in any given trial one of 41 potential target words. The analysis revealed a main effect of age (ß = 0.33, SE ß = 0.13, p = 0.01); and a significant estimate for the scaled average cross-linguistic frequency of phones in the target words (ß = 0.82, SE ß = 0.18, p \textless{} 0.001): Target words with phones found more frequently across languages had higher correct repetition scores, as shown in Figure \ref{fig:Fig2-xling-freq}. Averaging across items and participants, the Pearson correlation between scaled average cross-linguistic phone frequency and whole-item NWR scores was r(32) = 0.61.

@@

We next checked whether the association between whole-item NWR scores and cross-linguistic phone frequence could actually be due to frequency of the sounds within the language: One can suppose that sounds that occur more frequently across languages are also more frequent within a language, and therefore may be easier for children to represent and repeat. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings (Casillas et al., 2020) by counting the number of word types in which they occurred, and applied the natural logarithm.\footnote{We also carried out this analysis using token phone frequency, but this measure was not correlated with whole-item NWR scores, and therefore the fact that it did not explain away the predictive value of cross-linguistic phone frequency was less informative than the relationship discussed in the main text, with type frequencies.} Here, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Item-level average corpus-based phone frequencies were correlated with the item-level average cross-linguistic phone frequencies (r(27)=0.50, p \textless{} 0.01). Moreover, averaging across items and participants, the Pearson correlation between scaled average corpus phone frequency and whole-item NWR scores was r = 0.46. Therefore, we fit another mixed logistic regression, this time declaring as fixed effects both scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic phone frequency (ß = 0.82, SE ß = 0.27, p \textless{} 0.01) and age (ß = 0.33, SE ß = 0.13, p = 0.01) were significant predictors of whole-item NWR scores, but corpus phone frequency (ß = 0.01, SE ß = 0.25, p = 0.98) was not.

@@

Next, we investigated patterns of deletion and substitution (as insertion was not attested), this time looking at all attempts so as to base our generalizations on more data. Deletions were very rare: there were only 12 instances of deleted vowels (\textasciitilde{}0.28\% of all vowel targets), and 6 instances of deleted consonants (\textasciitilde{}0.19\% of all consonant targets). We therefore focus our qualitative description here on substitutions: There were 884 cases of substitutions, \textasciitilde{}17.83 of the 4957 phones found collapsing across all children and target words, so that substitutions constituted the frank majority of incorrect phones (\textasciitilde{}97.90 of unmatched phones).

Is it the case that cross-linguistically less common phones are more frequently substituted by more common ones than vice versa? We found that errors involving nasal vowel targets tended to result in a change in nasality (58 nasal targets produced as oral, \textasciitilde{}72\% of substitutions involving nasal vowels) more frequently than in a change in quality (23 nasal targets produced as a different nasal vowel, \textasciitilde{}28\% of substitutions involving nasal vowels). We found that errors involving nasal vowel targets tended to result in a change in nasality (52 oral targets produced as nasal, \textasciitilde{}20\% of substitutions involving nasal vowels) more frequently than in a change in quality (205 nasal targets produced as a different nasal vowel, \textasciitilde{}80\% of substitutions involving nasal vowels).

the proportion of errors for to be mistakenly produced as oral vowels than vice versa once prevalence in our stimuli was taken : 52 oral target vowels produced as nasal (\textasciitilde{}20\% of substituted oral target vowels) and Both oral and nasal vowels were also occasionally substituted by vowels of a different quality that preserved the target oral/nature airflow: 205 oral vowels (\textasciitilde{}2.16\% of targets and 23 nasal vowels (\textasciitilde{}2.36\% of targets).

Among consonants, complex consonants were occasionally substituted by simpler ones (specifically {[}ṭp{]}, {[}tp{]}, {[}kp{]}, {[}km{]}, {[}kṇ{]}, {[}mp{]}, {[}ɣ{]}, and {[}lβʲ{]} as {[}m{]}, {[}n{]}, {[}l{]}, {[}w{]}, {[}j{]}, {[}w{]}, {[}ṭ{]}, {[}g{]}, {[}p{]}, {[}t{]}, {[}k{]}, {[}f{]}, {[}h{]}, and {[}tʃ{]}), 250 times, in approximately 2.43\% of all complex consonant targets. Interestingly, simple consonants were sometimes also substituted by complex ones; this happened 2 times, in approximately 2.43\% of all simple consonant targets. Finally, simple consonants were mispronounced as other simple consonants (120 instances, \textasciitilde{}0.52\% of all simple consonant targets) and complex consonants as other complex consonants (55 instances, \textasciitilde{}0.98\% of all complex consonant targets). Note that the substitutions included phones that are not native to Yélî Dnye but do occur in English (e.g., {[}tʃ{]}); as these data come from careful transcriptions by a native Yélî Dnye speaker who is very fluent in English, we take these segmental substitutions faithfully as an indication that several of our participants have mastered production of some English phones, possibly produced within whole English word forms.

Finally, we looked into the frequency with which mispronunciations resulted in real words. In fact, two thirds of incorrect repetitions were recognizable as real words or phrases in Yélî Dnye or English: 63\%. This type of analysis is seldom reported. We could only find one comparison point: Castro-Caldas, Petersson, Reis, Stone-Elander, and Ingvar (1998) found that illiterate European Portuguese adults' NWR mispronunciations resulted in real words in 11.16\% of cases, whereas literate participants did so in only 1.71\% of cases. The percentage we observe here is much higher than reported in Castro and colleagues' study, but we do not know whether age, language, or even test structure explains this difference.

@@

Next, we inspected whether NWR scores varied as a function of word length (Table \ref{tab:tablength}). Participants scored much lower on monosyllables than on non-words of other lengths. This is likely due to the fact that the majority of monosyllables were designed to include sounds that are rare in the world's languages, which may indicate that they are hard to produce or perceive. Setting monosyllables aside, we observe the typical pattern of lower scores for longer items, although this is particularly salient for the whole-item scoring. While whole-item scoring is the most commonly reported NWR outcome, it is also the least forgiving. Averaging across participants and items, the Pearson correlation between length (2-4 syllables) and whole-item NWR scores was r(1) = -0.91. In a generalized binomial mixed model, we included 479 observations, from 40 children producing, in any given trial, one of 24 potential target words. The analysis revealed a positive effect of age (ß = 0.56, SE ß = 0.14, p \textless{} 0.001) and a negative but non-significant estimate for target length in number of syllables (ß = -0.15, SE ß = 0.33, p = 0.65).

@@

\caption{\label{tab:tablength}NWR measured in whole-word scores, and normalized Levenshtein Distance, separately for the four stimuli lengths.}

\begin{tabular}{ll}
\toprule
Word & \multicolumn{1}{c}{NLD}\\
\midrule
47 (22) & 41 (17)\\
79 (22) & 8 (9)\\
78 (19) & 7 (7)\\
74 (32) & 9 (12)\\
\bottomrule
\end{tabular}
